% v0.1
\chapter*{Conclusion} % chapter* je necislovana kapitola
\addcontentsline{toc}{chapter}{Conclusion} % rucne pridanie do obsahu
\markboth{Conclusion}{Conclusion} % vyriesenie hlaviciek

% TODO: 1 page
The importance of the bacteriophages as a research subject is rising mainly due to the decreasing effectiveness of antibiotic treatments.
Bacteriophages could be used as novel weapons in the fight against bacteria.
The goal of this work was to examine relationships between bacteriophage genomes and its bacterial host.

%They are already used to battle bacterial infections in the former Eastern bloc.
%In the countries like Georgia, cocktails of phages are produced and used to eliminate bacteria causing infection. They have been using this method for many years now.
%Unfortunately, due to the process by which these cocktails are produced makes the exact content of these cocktails unknown.
%This prevents western countries in adopting phage therapy as a treatment for bacterial infections.

At first, we collected publicly available data, annotated phage genomes for genes and clustered these genes according to their similarity.
Then, we proposed reduced representation of each bacteriophage based on genes, which were observed in its genome.
We demonstrated feasibility of this representation, using principal component analysis.
In representations that were transferred into lower dimensions we were able to roughly distinguish between clusters of bacteriophages with a distinct bacterial host.
Finally, we used a decision tree classifier to refine our prediction.
This machine learning method allowed us to create predictive models enabling us to assign a bacterial host with high accuracy using only information about the genome of the bacteriophage.
Models were also interpretable in the terms of gene clusters.
This allowed us to highlight clusters with high potential of being important for phage specificity.

The method demonstrated in this work could also be used for other biological traits, which depend on the genes included in phage genome.
This could bring a new insight on phages and their functions.
% We think, this method has a potential to help to overcome the barrier, which prevents use of bacteriophages in western countries.
